Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-04-18 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added description of how CHARMM36 72 lipid POPC and DPPC files were created.

{\it DPPC and POPC with 72 lipids}
\todo{Markus Miettinen, please describe how you made the files.}

Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

{\it POPC with 128 lipids}
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

{\it POPC with 512 lipids}
POPC The starting structures for simulations were constructed with
the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
used for the initial 0\% concentration and they were subsequently 
substituted by cholesterol molecules to fulfil the desired concentration.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
was used for both POPC and cholesterol. 
Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
were constrained with LINCS algorithm. Coordinates were written every 10 ps.
PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
method.

>>>>

{\it DPPC and POPC with 72 lipids.}
%
The starting structures in PDB format were downloaded from the
NIH/NHLBI Laboratory of Computational Biology Membrane Biophysics Section
website (\url{http://www.lobos.nih.gov/mbs/coords.shtml}),
which refers to these as the final structures (for DPPC after 40~ns, POPC after 35~ns)
of the NPT lipid bilayer trajectories used in the original CHARMM36 publication~\cite{klauda10}.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
%
The publicly available CHARMM36 parameters in Gromacs format
(September 2013 update: {\tt charmm36\_gmx\_format\_sep13.tgz})
from the MacKerell Lab website
(\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

{\it POPC with 128 lipids.}
%
The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website
(\url{http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html}).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters
(\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
was updated every 10 steps. Lennard-Jones interactions were switched to zero
between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.

{\it POPC with 512 lipids.}
%
POPC The starting structures for simulations were constructed with
the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
used for the initial 0\% concentration and they were subsequently 
substituted by cholesterol molecules to fulfil the desired concentration.
The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
was used for both POPC and cholesterol. 
Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
were constrained with LINCS algorithm. Coordinates were written every 10 ps.
PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
method.
